Necessity is the mother of genetic invention

### 2017-11-03 ms.tex
cite Fisher for reproductive value; fix section referencing - was messed by hyperref+titlesec so just switched subsections to paragraph

\usepackage{amssymb,amsmath,amsthm}
\usepackage{longtable,booktabs}
\PassOptionsToPackage{hyphens}{url} % url is loaded by hyperref
\usepackage[unicode=true]{hyperref}
\usepackage[auth-sc]{authblk}
\usepackage{titlesec}

\titleformat{\subsection}[runin]
  {\bf}{\thesection}{1em}{}

@@

\subsection*{General model.}
Consider a large haploid population with an arbitrary (possibly infinite) set of types $A_1, A_2, \ldots$.
The change in the frequencies of these types is described by the transformation $f \to f'$: 
\begin{equation}\label{eq:model}
\bar{\omega} f' = (I-C+MC)D f,
\end{equation}
where $f_k$ is the frequency of type $A_k$ ($f=(f_1, f_2, \ldots)$ is a stochastic vector: $\sum_k{f_k} = 1$);
$C_k$ is the probability that an individual of type $A_k$ will transition to another type ($C$ is a diagonal non-negative matrix with diagonal elements $C_k$);
$M_{i,j}$ is the transition probability from type $A_j$ to type $A_i$ given that an individual of type $A_j$ is transitioning ($M$ is a column-stochastic matrix: $\sum_j {M_{i,j}} = 1$ for any column \emph{j}); 
$D_k$ is the relative fitness of type $A_k$ ($D$ is a diagonal matrix with diagonal elements $D_k$, where we can assume $0 < D_k \le 1$, i.e., fitness values are relative to the fitness of the fittest type); 
$I$ is the identity matrix;
and $\bar{\omega}$ is a normalizing factor such that $\sum_k{f'_k}=1$ and is equal to the population mean fitness (see~\nameref{sec:AppA}).

@@

\subsection*{Mutation modification model 1.}\label{sec:mut_model_1}

@@

\subsection*{Mutation modification model 2.}\label{sec:mut_model_2}

@@

\subsection*{Migration modification model.}\label{sec:mig_model_1}

@@

\begin{equation}
\begin{aligned}
\frac{\partial \bar{\omega}^*}{\partial C_k} = 
v_k D_k (u\tr[M]_k - u_k) = \\
D_k (\sum_i{u_i M_{i,k} v_k} - u_k v_k).
\end{aligned}
\end{equation}
Here, $u_k$ can be regarded as the \emph{reproductive value} of type $k$ (see~\nameref{sec:AppC}), which gives the relative contribution of type $k$ to the long-term population.
Consequently, $u_k v_k$ is the \emph{ancestor frequency} \citep{Hermisson2002} of type $k$, which is the fraction of the equilibrium population descended from type $k$.
The sum $\sum_i{u_i M_{i,k} v_k}$ can be similarly interpreted as the fraction of the equilibrium population descended from individuals that transitioned from type $k$ to another type (via the $k$ column of the transition matrix $M$).

@@

\section*{Appendices}
% Appendix A: mean fitness
\subsection*{Appendix A}\label{sec:AppA}
\textbf{TODO} reference to Fisher 1930 or Fisher 1958 (Caswell 1978 cites the latter in pg. 217 just above eq. 4).

@@

% Appendix C: Fisher's reproductive value
\subsection*{Appendix C}\label{sec:AppC}
\begin{frv}
Let $M$ be a stochastic matrix and $D$ be a diagonal matrix with positive diagonal elements. 
The elements of the left \emph{Perron} eigenvector $u$ of the matrix $MD$ can be regarded as \emph{Fisher's reproductive values}~\citep{Fisher1930,Grafen2006}.
\end{frv}

\emph{Fisher's reproductive values} can be understood as follows.
Consider the dynamics not of frequencies but of absolute population sizes such that the vector of the number of individuals of each type at time $t$ is $n(t)$ and $f_k(t) = n_k(t) / \sum_i{n_i(t)}$.
The dynamics are
\begin{equation}
n(t) = (MD)^t n(0).
\end{equation}

@@

Now take the long termlimit $t \goesto \infty$.
By assumption, $MD$ is irreducible, so $\lambda_i < \lambda_1$ for all $i > 1$.
Therefore,
\begin{equation}
\lim_{t \goesto \infty} \Big( \frac{\lambda_k}{\lambda_1} \Big)^t = 0
\end{equation}
for all $k > 1$, and
\begin{equation}
\lim_{t \goesto \infty}\frac{N(j,t)}{N(k,t)} =
\frac{e\tr V (e_1 e_1\tr) U\tr e_j} { e\tr V (e_1 e_1\tr) U\tr e_k} = 
\frac{ e\tr v u_j} { e\tr v  u_k} =
\frac{u_j} {u_k}.
\end{equation}

>>>>

\usepackage{amssymb,amsmath,amsthm}
\usepackage{longtable}
\usepackage[hyphens]{url}
\PassOptionsToPackage{hyphens}{url} % url is loaded by hyperref
\usepackage[unicode=true]{hyperref}
%\usepackage{nameref} % included in hyperref
\usepackage[auth-sc]{authblk}

% messes up nameref
%\usepackage{titlesec}
%\titleformat{\subsection}[runin]
%  {\bf}{\thesection}{1em}{}

@@

\paragraph*{General model.}
Consider a large haploid population with an arbitrary (possibly infinite) set of types $A_1, A_2, \ldots$.
The change in the frequencies of these types is described by the transformation $f \to f'$: 
\begin{equation}\label{eq:model}
\bar{\omega} f' = (I-C+MC)D f,
\end{equation}
where $f_k$ is the frequency of type $A_k$ ($f=(f_1, f_2, \ldots)$ is a stochastic vector: $\sum_k{f_k} = 1$);
$C_k$ is the probability that an individual of type $A_k$ will transition to another type ($C$ is a diagonal non-negative matrix with diagonal elements $C_k$);
$M_{i,j}$ is the transition probability from type $A_j$ to type $A_i$ given that an individual of type $A_j$ is transitioning ($M$ is a column-stochastic matrix: $\sum_j {M_{i,j}} = 1$ for any column \emph{j}); 
$D_k$ is the relative fitness of type $A_k$ ($D$ is a diagonal matrix with diagonal elements $D_k$, where we can assume $0 < D_k \le 1$, i.e., fitness values are relative to the fitness of the fittest type); 
$I$ is the identity matrix;
and $\bar{\omega}$ is a normalizing factor such that $\sum_k{f'_k}=1$ and is equal to the population mean fitness (see~\nameref{sec:AppA}).

@@

\paragraph*{Mutation modification model 1.}\label{sec:mut_model_1}

@@

\paragraph*{Mutation modification model 2.}\label{sec:mut_model_2}

@@

\paragraph*{Migration modification model.}\label{sec:mig_model_1}

@@

\begin{equation}
\begin{aligned}
\frac{\partial \bar{\omega}^*}{\partial C_k} = 
v_k D_k (u\tr[M]_k - u_k) = \\
D_k (\sum_i{u_i M_{i,k} v_k} - u_k v_k).
\end{aligned}
\end{equation}
Here, $u_k$ can be regarded as the \emph{reproductive value}~\citep[pg.~27]{Fisher1930} of type $k$ (see~\nameref{sec:AppC}), which gives the relative contribution of type $k$ to the long-term population.
Consequently, $u_k v_k$ is the \emph{ancestor frequency} \citep{Hermisson2002} of type $k$, which is the fraction of the equilibrium population descended from type $k$.
The sum $\sum_i{u_i M_{i,k} v_k}$ can be similarly interpreted as the fraction of the equilibrium population descended from individuals that transitioned from type $k$ to another type (via the $k$ column of the transition matrix $M$).

@@

\section*{Appendices}
% Appendix A: mean fitness
\subsection*{Appendix A}\label{sec:AppA}

@@

% Appendix C: Fisher's reproductive value
\subsection*{Appendix C}\label{sec:AppC}
\begin{frv}
Let $M$ be a stochastic matrix and $D$ be a diagonal matrix with positive diagonal elements. 
The elements of the left \emph{Perron} eigenvector $u$ of the matrix $MD$ can be regarded as \emph{Fisher's reproductive values}~\citep[pg.~27]{Fisher1930}
\end{frv}

\emph{Fisher's reproductive values} can be understood as follows~\citep{Grafen2006}.
Consider the dynamics not of frequencies but of absolute population sizes such that the vector of the number of individuals of each type at time $t$ is $n(t)$ and $f_k(t) = n_k(t) / \sum_i{n_i(t)}$.
The dynamics are
\begin{equation}
n(t) = (MD)^t n(0).
\end{equation}

@@

Now take the limit $t \goesto \infty$.
By assumption, $MD$ is irreducible, so $\lambda_i < \lambda_1$ for all $i > 1$.
Therefore,
\begin{equation}
\lim_{t \goesto \infty} \Big( \frac{\lambda_k}{\lambda_1} \Big)^t = 0
\end{equation}
for all $k > 1$, and
\begin{equation}
\lim_{t \goesto \infty}\frac{N(j,t)}{N(k,t)} =
\frac{e\tr V (e_1 e_1\tr) U\tr e_j} { e\tr V (e_1 e_1\tr) U\tr e_k} = 
\frac{ e\tr v u_j} { e\tr v  u_k} =
\frac{u_j} {u_k}.
\end{equation}
